Extended Data

### 2022-05-07 doc/submissions/2022_04_Nature/Response to reviews.Rmd

**We agree with this comment and specify our motivation in the methods section: "We defined CTW for *DOY~25~*, as opposed to other parameters describing the timing of growth, because spring temperatures should have the most direct influence on this variable."**

**We also repeat the analyses using a definition of CTW that is independent of *DOY~25~* and all other response variables. As explained in the methods, "To ensure that patterns were robust under an alternative definition of CTW, and to parallel the monthly time windows used in our tree-ring analysis (detailed below; Fig. 3, Extended Data Figure 6-7), we also ran analyses where we fixed the CTW identified by climwin to be the month with the most days in the CTW (Extended Data Table 2) each critical window."**

I appreciated the clarifications done concerning the data cleaning process. However, I didn’t find in the manuscript the quantity of data that have been removed at it stages. I believe it is worth mentioning it.

*Our data cleaning methods were split into two sections. First, we performed manual inspection of the raw dendrometer band measurements to remove unreliable records. Second, we removed models which returned parameters that we determined to be outliers. We provide quanitity of data removed for the second section, but not the first. When we revisited the first section to provide numbers, we noticed that the manually cleaned dendrometer band measurements we originally planned to feed into the modeling process had been replaced by the raw dendrometer band measurements with no manual cleaning done. As such, we have decided to remove the text, "The raw dendrometer band data were screened before analysis to remove records or entire tree-years that were either unreliable or inappropriate for our analysis. First, we manually screened the data for three classes of errors. First, when a measurement was drastically different from previous and following measurements, it was assumed to be a human error and removed. Second, when measurements remained essentially unchanged for several readings, followed by a sudden jump then return to a normal growth pattern, this was assumed to be a case where the band was stuck on the tree bark and then released. In these cases, the full annual record for the tree was removed. Third, data points that deviated substantially from normal growth patterns, but for unknown causes, were removed. If a majority of the data points fell into this class within a tree-year, the entire year was removed from the analysis." from the methods section. *

*We believe that losing this step in the data cleaning process does not significantly alter the quantity or quality of the data we ultimately used for analysis. This is because any tree-year which may have been removed using the criteria above would either fail the model fitting process, or would be removed if it produced an outlier model. Meaning, since there were not a substantial number of tree-years where manually cleaning was necessary, any model that was potentially compromised by faulty measurements would be removed by our post-model cleaning method.*

### 2022-05-07 doc/manuscript/Dow_manuscript_Nature.Rmd
draft response
@teixeirak let me know what you think

The raw dendrometer band data were screened before analysis to remove records or entire tree-years that were either unreliable or inappropriate for our analysis. 
First, we manually screened the data for three classes of errors. 
First, when a measurement was drastically different from previous and following measurements, it was assumed to be a human error and removed.
Second, when measurements remained essentially unchanged for several readings, followed by a sudden jump then return to a normal growth pattern, this was assumed to be a case where the band was stuck on the tree bark and then released. 
In these cases, the full annual record for the tree was removed.
Third, data points that deviated substantially from normal growth patterns, but for unknown causes, were removed. 
If a majority of the data points fell into this class within a tree-year, the entire year was removed from the analysis.
We also removed tree-years with small or negligible total growth ($\Delta DBH$ $\le 0.005$ cm; SCBI = 26, Harvard Forest = 253), and tree-years where the first intra-annual measurement was later than the first spring survey (trees that were missed in the initial census; SCBI = 22, Harvard Forest = 8).
These were removed because they reduce the reliability of predicted growth in the modeled curves.

>>>>

We removed tree-years with small or negligible total growth ($\Delta DBH$ $\le 0.005$ cm; SCBI = 26, Harvard Forest = 253), and tree-years where the first intra-annual measurement was later than the first spring survey (trees that were missed in the initial census; SCBI = 22, Harvard Forest = 8).
These were removed because they reduce the reliability of predicted growth in the modeled curves.
